\myTitle

### 2024-01-11 body-text-of-paper.tex
Added newhybs figure.

{\sc hybridlab} was conceived at a time when
the data sets used in molecular ecology universally included very few markers, such that
it was reasonable to assume the markers were unlinked, and, hence, independently segregating.
Today, however, with data sets of thousands of markers, that assumption is badly violated---alleles
on the same chromosome are likely to be segregated together into the next generation.
Simulating linked loci as independent creates simulated hybrids, beyond the $F1$ category, with 
artificially low variance in admixture fraction, leading to the false interpretation that different
hybrid categories are more easily distinguished than they are. Therefore, in evaluating descendants of migrants within a simulated framework, it is imperative to take into consideration the pattern of
linkage among loci available for analysis. This was done recently to assess the power of a SNP panel for identifying descendants of escaped, farm-raised Atlantic salmon \citep{wringe2019development}; however the method used (described at https://eriqande.github.io/SalarHybPower/) was developed in an {\em ad hoc} fashion and was not easily extended to
complex hybridization scenarios, and was not at all user-friendly.

@@

@@

Genetic material is segregated through
a GSP using these steps, done upon each node in an order such that the steps
are run on each node's parents before being run on the node itself:
\begin{enumerate}
\item Each genome in a triangle above a founder node represents a gamete
carrying a single, haploid copy of the genome.
These gametes are united within the founder individual and then
recombination occurs between each chromosome
with probability $\frac{1}{2}$ and within each chromosome according to
a user-specified recombination map to create two new gametes, which are
recombined version of the original ones.  These two gametes are then segregated at random down the edge (or edges) from the founder node. Upon each edge, $i$ from founder node $f$, $G^-_{f,i}$ gametes are segregated. 
\item Each non-founder node, $x$, will have exactly two edges coming
into it---one from each parent.  Each of the $G^+_{x,1}$ gametes coming in from the
first edge is united with a single one of the $G^+_{x,2}$ gametes coming
in on the second edge (in a random order).
\item If a non-founder node $x$ has an edge to a sample node,
a randomly chosen $S_x$ of the united pairs of gametes are delivered
to the sample node to constitute the $S_x$ samples taken from individual
node $x$. Any remaining united pairs of gametes are treated as in (4) below.
\item United pairs of gametes (including those remaining after delivery
to a sample node) in each non-founder node $x$ are segregated into
gametes with recombination, and each of these gametes is assigned, in
random order, to the gamete edges proceeding downward out of the node,
with the number of gametes assigned to the $i\thh$ downward edge being
$G^-_{x,i}$.
\end{enumerate}
At the end of this process, the samples simulated (as united pairs of gametes)
into the sample nodes will contain all of the genetic material (and no more) found among the founders, but that
material will have been segregated and recombined according to the pedigree.  Thus the
samples obtained from a GSP can be simulated with the expected admixture fractions
for any hybrid category that can be specified by a pedigree.  They have also been simulated
in a way that 1) reflects physical linkage---each individual inherits segments of the genome
simulated using a recombination map; 2) does not duplicate any genetic material---the sampling of
genetic material is entirely without replacement so it will not induce RISPI; and 3) no genetic material has been lost---all the genetic material of the founders
is represented in the samples, maximizing the number of simulated individuals and markers available to
use in downstream analyses.

@@

Genotypes for

@@

Boing.

@@

\section*{Acknowledgements}
Research was supported in part by the USDA-APHIS-National Feral Swine Damage
Management Program. We are grateful for the numerous Wildlife Services and Missouri
Department of Conservation employees that contributed to genetic sampling of wild pigs
within southeastern Missouri. Writing of this paper was facilitated by the Mobile High Altitude Venue for Ecological Analysis,
Genetics and Statistics, through which ECA, MGD, and TJS were scientists-in-residence on location in
Cheyenne, WY for 3 days in October 2023. This is contribution number mHAVEAGAS-003.
This work utilized the Alpine high performance computing resource at the University of Colorado Boulder.
Alpine is jointly funded by the University of Colorado Boulder, the University of Colorado Anschutz, Colorado State University,
and the National Science Foundation (award 2201538).

>>>>

{\sc hybridlab} was conceived at a time when
the data sets used in molecular ecology universally included very few markers, such that
it was reasonable to assume the markers were unlinked, and, hence, independently segregating.
Today, however, with data sets of thousands of markers, that assumption is badly violated---alleles
on the same chromosome are likely to be segregated together into the next generation.
Simulating linked loci as independent creates simulated hybrids, beyond the $F1$ category, with 
artificially low variance in admixture fraction, leading to the false interpretation that different
hybrid categories are more easily distinguished than they are. Therefore, in evaluating descendants of migrants within a simulated framework, it is imperative to take into consideration the pattern of
linkage among loci available for analysis. This was done recently to assess the power of a SNP panel for identifying descendants of escaped, farm-raised Atlantic salmon \citep{wringe2019development}; however the method used (described at https://eriqande.github.io/SalarHybPower/) was developed in an {\em ad hoc} fashion, was not easily extended to
complex hybridization scenarios, and was not at all user-friendly.

@@

NOTE, WE ALSO DID THESE SIMULATIONS USING GSCRAMBLE AND WE SHOULD NOTE
HOW THOSE CAME OUT.

\subsection*{Physical Linkage Matters}

ECA STILL NEEDS TO DESCRIBE AND INCLUDE THE SIMULATIONS THAT HE DID ON THIS.
THAT WORK IS ALL DONE, I JUST NEED TO PUT IT IN HERE, ALONG WITH THE FIGURE.

@@

Genetic material is segregated through
a GSP using these steps, done upon each node in an order such that the steps
are run on each node's parents before being run on the node itself:
\begin{enumerate}
\item Each genome in a triangle above a founder node represents a gamete
carrying a single, haploid copy of the genome.
These gametes are united within the founder individual and then
recombination occurs between each chromosome
with probability $\frac{1}{2}$ and within each chromosome according to
a user-specified recombination map to create two new gametes, which are
recombined versions of the original ones.  These two gametes are then segregated at random down the edge (or edges) from the founder node. Upon each edge, $i$ from founder node $f$, $G^-_{f,i}$ gametes are segregated. 
\item Each non-founder node, $x$, will have exactly two edges coming
into it---one from each parent.  Each of the $G^+_{x,1}$ gametes coming in from the
first edge is united with a single one of the $G^+_{x,2}$ gametes coming
in on the second edge (in a random order).
\item If a non-founder node $x$ has an edge to a sample node,
a randomly chosen $S_x$ of the united pairs of gametes are delivered
to the sample node to constitute the $S_x$ samples taken from individual
node $x$. Any remaining united pairs of gametes are treated as in (4) below.
\item United pairs of gametes (including those remaining after delivery
to a sample node) in each non-founder node $x$ are segregated into
gametes with recombination, and each of these gametes is assigned, in
random order, to the gamete edges proceeding downward out of the node,
with the number of gametes assigned to the $i\thh$ downward edge being
$G^-_{x,i}$.
\end{enumerate}
At the end of this process, the samples simulated (as united pairs of gametes)
into the sample nodes will contain all of the genetic material (and no more) found among the founders, but that
material will have been segregated and recombined according to the pedigree.  Thus the
samples obtained from a GSP can be simulated with the expected admixture fractions
for any hybrid category that can be specified by a pedigree.  They have also been simulated
in a way that 1) reflects physical linkage---each individual inherits segments of the genome
simulated using a recombination map; 2) does not duplicate any genetic material---the sampling of
genetic material is entirely without replacement so it will not induce RISPI; and 3) no genetic material has been lost---all the genetic material of the founders
is represented in the samples, maximizing the number of simulated individuals and markers available to
use in downstream analyses.

@@

We used the archived data from \citet{rizza2023limited}, drawing upon their 65 markers
that are fixed or are nearly fixed for alternate alleles between rainbow trout and cutthroat trout.
They designated 634 fish as pure cutthroat trout and 213 as pure steelhead.  We used the
genotypes of these individuals as input to \gscramble{}.

\citeauthor{rizza2023limited} mapped the 65 markers to the Omyk\_v1.0 genome assembly (RefSeq 
Accession: GCF\_002163495.1), so we used
the linkage map described in \citep{pearse2019sex} that was prepared by linkage mapping 
of $\approx$57,000 markers described in \citet{palti2015development} to Omyk\_v1.0. We mapped the $\approx$57,000 markers to Omyk\_v1.0, and then resolved inconsistencies to produce a map
of 32,923 markers throughout the genome and rendered it to plink format for use with \gscramble{}.

In the original paper, \citet{rizza2023limited} identified 15 $F_1$s,
1 $F_2$,
6 cutthroat backcrosses ($BC1$s) and 
6 second-generation cutthroat backcrosses ($BC2$s),
and 1 fish of an unknown hybrid category that may have been the
product of a $BC2$ crossed with a $BC1$.   In our simulations, we sought to simulate
a similar collection of hybrid individuals, so, for each replicate, we used
\gscramble{} to simulate 16 $F_1$s, 4 $F_2$s, 6 $BC1$s, 4 $BC2$s, 4 $BC3$s,  4 individuals
that were the products of a $BC1$ mating with a $BC2$, and finally, 4 individuals that were simulated pure cutthroat. Individuals not used as founders
in \gscramble{} were retained as pure cutthroat and pure steelhead.  The goal of the simulation
was to assess how well these different hybrid categories can be distinguished from one another
when using either all 65 markers, or just the 30 markers occurring on different chromosome arms
used by  \citet{rizza2023limited}.

HEY! I SHOULD ALSO DO A SET OF SIMULATIONS IN WHICH THE 65 MARKERS ARE
SIMULATED AS IF THEY ARE INDEPENDENTLY SEGREGATING---I THINK THAT MEANS
ASSIGNING EACH TO ITS OWN CHROMOSOME.  BUT MAYBE WORRY ABOUT THAT
LATER.

We performed 500 replicate simulations with \gscramble{} using either 30 or 65 markers.  Each
simulated data set was then analyzed using {\sc NewHybrids} \citep{anderson2002model}. Individuals not used as founders were given the ``s'' and ``z'' flags, indicating that they were considered training samples.  The remaining 46 individuals were treated as possible hybrids and their genetic data used to calculate the posterior probability that they belonged to one of 10 hybrid categories as shown in Table~\ref{tab:newhybcats}, using 25,000 MCMC sweeps and discarding the first 5,000 as burn-in.
%%%%%%%%%%%%
\begin{table}
\caption{The 10 hybrid categories considered for inference with {\sc NewHybrids} and the
expected frequency of their genotypes having 2 ($f(C,C)$), 1 ($f(C,S)$) or 0 ($f(S,S)$) copies of ancestry from cutthroat trout.}
\label{tab:newhybcats}
{\small
\begin{tabular*}{\columnwidth}{@{\extracolsep{\fill}} lrrr}
\hline\hline
Category	& $f(C,C)$ &	$f(C,S)$ & 	$f(S,S)$ \\ \hline
\input{inputs/newhybs-cats.tsv}
\end{tabular*}
}
\vspace*{-2.3ex}\hrule\vspace*{0.3ex}\hrule
\end{table}
%%%%%%%%%%%%

@@

PRETTY MUCH DONE HERE EXCEPT I WANT TO SIMULATE THINGS
UNLINKED.
%%%%%%%%%%
\begin{figure}
\newcommand{\nhbpcap}{\footnotesize Gonna need a caption}
\includegraphics[width=\columnwidth]{figures/newhybs-texed-crop.pdf}
\caption[\nhbpcap]{\nhbpcap}
\label{fig:newhyb-sims-barplot}
\end{figure}
%%%%%%%%%%

@@

\section*{Acknowledgements}
Research was supported in part by the USDA-APHIS-National Feral Swine Damage
Management Program. We are grateful for the numerous Wildlife Services and Missouri
Department of Conservation employees that contributed to genetic sampling of wild pigs
within southeastern Missouri. We thank Yniv Palti for providing flanking sequences and the linkage map for 57K markers for making a recombination map for the cutthroat/rainbow trout example. Writing of this paper was facilitated by the Mobile High Altitude Venue for Ecological Analysis,
Genetics and Statistics, through which ECA, MGD, and TJS were scientists-in-residence on location in
Cheyenne, WY for 3 days in October 2023. This is contribution number mHAVEAGAS-003.
This work utilized the Alpine high performance computing resource at the University of Colorado Boulder.
Alpine is jointly funded by the University of Colorado Boulder, the University of Colorado Anschutz, Colorado State University,
and the National Science Foundation (award 2201538).
